---
title: "Exercixe 3 2540122836"
author: "Simen Ferdinand Djamhari"
output:
  html_document:
    df_print: paged
---


```{r}
#source('create_datasets.R')

library(readr)
library(dplyr)
library(ggplot2)
library(openintro)

cars <- read.csv("https://assets.datacamp.com/production/course_1796/datasets/cars04.csv")
comics <- read.csv("https://assets.datacamp.com/production/course_1796/datasets/comics.csv")
life <- read.csv("https://assets.datacamp.com/production/course_1796/datasets/life_exp_raw.csv")
```


```{r}
# Print the first rows of the data
head(comics)
```

__EXPLANATION__:
1. comics data set have 11 column (name, id, align, eye, hair, gender, gsm, alive, appearances, first_appear, and publisher)

```{r}
# Check levels of align
comics$align <- as.factor(comics$align)
levels(comics$align)
```
__EXPLANATION__:
1. There's 4 align in comics (Bad, Good, Neutral, Reformed Criminal)
```{r}
# Check the levels of gender
comics$gender <- as.factor(comics$gender)
levels(comics$gender)
```
__EXPLANATION__:
1. There's 3 gender in comics (Female, Male, Other)

```{r}
# Create a 2-way contingency table
table(comics$align, comics$gender)
```
__EXPLANATION__:
1. As we can see from the table, there are overwhelmingly more Male than Female in comics. The highest is "Bad" "Male" that has 7561 frequency

```{r}
# Print tab
tab <- table(comics$align, comics$gender)
tab
```
##Dropping levels
```{r}
# Remove align level
comics <- comics %>%
  filter(align != 'Reformed Criminals') %>%
  droplevels()

levels(comics$align)
```
__EXPLANATION__:
1. Because Reformed Criminals is not that useful, it is decided to get drop.

##Side-by-side barcharts
```{r}
# Create side-by-side barchart of gender by alignment
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "dodge")
```
```{r}
# Create side-by-side barchart of alignment by gender
ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(positio = "dodge") +
  theme(axis.text.x = element_text(angle = 90))
```

__EXPLANATION__:
1. The comics dataset has more males than female ,other ,and NA
2. Male has more on Bad, Good, and Neutral align than female ,other ,and NA


##Counts vs. proportions
```{r}
# simplify display format
options(scipen = 999, digits = 3) 

## create table of counts
tbl_cnt <- table(comics$id, comics$align)
tbl_cnt
```
```{r}
# Proportional table
# All values add up to 1
prop.table(tbl_cnt)
```
__EXPLANATION__:
1. Because "Bad" "Secret" has the most count(4493), it also has the most propotion(0.289609)

```{r}
sum(prop.table(tbl_cnt))
```

```{r}
# All rows add up to 1
prop.table(tbl_cnt, 1)
```
__EXPLANATION__:
1. prop.table(tbl_cnt, 1) function make the proportion between each row, that's why if we see the Bad Unknown, it has the most proportion. It's because the proportion table compare it betwwen it's row. 
```{r}
# Coluns add up to 1
prop.table(tbl_cnt, 2)
```
__EXPLANATION__:
1. prop.table(tbl_cnt, 1) function make the proportion between each row, while prop.table(tbl_cnt, 2) make the proportion between each column. The Secret has the most proportion from Bad while Public has the most proportion on Good. 



```{r}
ggplot(comics, aes(x = id, fill = align)) + 
  geom_bar(position = "fill") + 
  ylab("proportion")
```
```{r}
ggplot(comics, aes(x = align, fill = id)) + 
  geom_bar(position = "fill") + 
  ylab("proportion")
```

__EXPLANATION__:
1. There are no Good align that is Unknown
2. Most Good align come from public
3. Most bad align come from Secret

##Conditional proportions
```{r}
tab <- table(comics$align, comics$gender)
options(scipen = 999, digits = 3) # Print fewer digits
prop.table(tab)     # Joint proportions
```
```{r}
prop.table(tab, 2)
```

## Counts vs. proportions (2)
```{r}
# Plot of gender by align
ggplot(comics, aes(x = align, fill = gender)) +
  geom_bar()
```
```{r}
# Plot proportion of gender, conditional on align
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "fill")
```

__EXPLANATION__:
1. proportionally, there are more Bad Male rather than Bad Female and Bad Other
2. proportionally, there are more Good Female rahter than Bad Male and Bad Other


##Distribution of one variable
```{r}
table(comics$id)
```

```{r}
# Simple barchart
ggplot(comics, aes(x = id)) + geom_bar()
```
__EXPLANATION__:
1. Secret has the most count on id

```{r}
ggplot(comics, aes(x = id)) + 
  geom_bar() + 
  facet_wrap(~align)
```
__EXPLANATION__:
1. Secret has the most count on Bad
2. Public has the most count on Good and Neutral

##Marginal barchart

```{r}
# Change the order of the levels in align
comics$align <- factor(comics$align, 
                       levels = c("Bad", "Neutral", "Good"))

# Create plot of align
ggplot(comics, aes(x = align)) + 
  geom_bar()
```

##Conditional barchart
```{r}
# Plot of alignment broken down by gender
ggplot(comics, aes(x = align)) + 
  geom_bar() + 
  facet_wrap(~gender)
```
__EXPLANATION__:
1. Most Male and NA in comics is Bad
2. Most Female in comics dataset is Good

##Improve piechart
```{r}
pies <- data.frame(flavors = as.factor(rep(c("apple", "blueberry", "boston creme", "cherry", "key lime", "pumpkin", "strawberry"), times = c(17, 14, 15, 13, 16, 12, 11))))

# Put levels of flavor in descending order
lev <- c("apple", "key lime", "boston creme", "blueberry", "cherry", "pumpkin", "strawberry")
pies$flavor <- factor(pies$flavor, levels = lev)

# Create barchart of flavor
ggplot(pies, aes(x = flavor)) + 
  geom_bar(fill = "chartreuse") + 
  theme(axis.text.x = element_text(angle = 90))
```
__EXPLANATION__:
1. Apples is the highest in flavor while strawberry is the lowest

##Exploring Numerical Data
```{r}
# A dot plot shows all the datapoints
ggplot(cars, aes(x = weight)) + 
  geom_dotplot(dotsize = 0.4)
```

```{r}
# A histogram groups the points into bins so it does not get overwhelming
ggplot(cars, aes(x = weight)) + 
  geom_histogram(dotsize = 0.4, binwidth = 500)
```
```{r}
# A density plot gives a bigger picture representation of the distribution
# It more helpful when there is a lot of data
ggplot(cars, aes(x = weight)) + 
  geom_density()
```
```{r}
# A boxplot is a good way to just show the summary info of the distriubtion
ggplot(cars, aes(x = 1, y = weight)) + 
  geom_boxplot() +
  coord_flip()
```

__EXPLANATION__:
1. The weight in cars dataset in concentrated at 3000 - 4000 range


## Faceted histogram
```{r}
# Load package
library(ggplot2)

# Learn data structure
str(cars)
```

```{r}
# Create faceted histogram
ggplot(cars, aes(x = city_mpg)) +
  geom_histogram() +
  facet_wrap(~ suv)
```



##Boxplots and density plots
```{r}
unique(cars$ncyl)
```
```{r}
table(cars$ncyl)
```
```{r}
# Filter cars with 4, 6, 8 cylinders
common_cyl <- filter(cars, ncyl %in% c(4,6,8))

# Create box plots of city mpg by ncyl
ggplot(common_cyl, aes(x = as.factor(ncyl), y = city_mpg)) +
  geom_boxplot()
```
__EXPLANATION__:
1. 4 has the highest amount of count in city_cars

```{r}
# Create overlaid density plots for same data
ggplot(common_cyl, aes(x = city_mpg, fill = as.factor(ncyl))) +
  geom_density(alpha = .3)
```

__EXPLANATION__:
1. 8 Has the highest density at 18, while 6 at 19 - 21, and 4 at 23- 25


##Distribution of one variable
#Marginal and conditional histograms
```{r}
# Create hist of horsepwr
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram() +
  ggtitle("Horsepower distribution")
```
```{r}
# Create hist of horsepwr for affordable cars
cars %>% 
  filter(msrp < 25000) %>%
  ggplot(aes(horsepwr)) +
  geom_histogram() +
  xlim(c(90, 550)) +
  ggtitle("Horsepower distribtion for msrp < 25000")
```

__EXPLANATION__:
1. The highest horsepower car in the less expensive range has just under 250 horsepower.

#Three binwidths
```{r}
# Create hist of horsepwr with binwidth of 3
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 3) +
  ggtitle("binwidth = 3")
```
```{r}
# Create hist of horsepwr with binwidth of 30
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 30) +
  ggtitle("binwidth = 30")
```
```{r}
# Create hist of horsepwr with binwidth of 60
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 60) +
  ggtitle("binwidth = 60")
```

__EXPLANATION__:
1. From the first graph, it's a little bit hard to see where the data is concentrated at. That's wny we make the binwidth bigger to see it more clearly. From the 30 binwidth, the data concentrated at 195 - 210 whole from 60 binwidth, the data concentrated at 150 - 210

##Box plots
#Box plots for outliers
```{r}
# Construct box plot of msrp
cars %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
```
```{r}
# Exclude outliers from data
cars_no_out <- cars %>%
  filter(msrp < 100000)

# Construct box plot of msrp using the reduced dataset
cars_no_out %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
```
__EXPLANATION__:
1. The outliers is above 62500

# Plot selection
```{r}
# Create plot of city_mpg
cars %>%
  ggplot(aes(x = 1, y = city_mpg)) +
  geom_boxplot()
```
```{r}
cars %>%
  ggplot(aes(city_mpg)) +
  geom_density()
```
__EXPLANATION__:
1. For city_mpg, The outliers is above 29
2. The data is concentrated at around 18 - 19

```{r}
# Create plot of width
cars %>%
  ggplot(aes(x = 1, y = width)) +
  geom_boxplot()
```
```{r}
cars %>%
  ggplot(aes(x = width)) +
  geom_density()
```
__EXPLANATION__:
1. For width, The outliers is above 78
2. The data is concentrated at around 68 - 72

##Visualization in higher dimensions
# 3 variable plot
```{r}
# Facet hists using hwy mileage and ncyl
common_cyl %>%
  ggplot(aes(x = hwy_mpg)) +
  geom_histogram() +
  facet_grid(ncyl ~ suv) +
  ggtitle("hwy_mpg by ncyl and suv")
```
__EXPLANATION__:
1. Across both SUVs and non-SUVs, mileage tends to decrease as the number of cylinders increases.

##Numerical Summaries
#Measures of center
```{r}
head(life)
```
__EXPLANATION__:
1. life dataset has 6 column (State, Country, fips, Year, Female.life.expectancy..years., Female.life.expectancy..national..years.)

```{r}
x <- head(round(life$Female.life.expectancy..years.), 11)
x
```
```{r}
sum(x)/11
mean(x)

sort(x)
median(x)

table(x)
```
__EXPLANATION__:
1. The Mean from the first 11 Female.life.expectancy..years. in life dataset is 77.2
2. The Median from the first 11 Female.life.expectancy..years. in life dataset is 77
3. The Mode from the first 11 Female.life.expectancy..years. in life dataset is 77

```{r}
library(gapminder)
str(gapminder)
```
```{r}
# Create dataset of 2007 data
gap2007 <- filter(gapminder, year == 2007)

# Compute groupwise mean and median lifeExp
gap2007 %>%
  group_by(continent) %>%
  summarize(mean(lifeExp),
            median(lifeExp))
```
```{r}
# Generate box plots of lifeExp for each continent
gap2007 %>%
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```
__EXPLANATION__:
1. Oceania Continent has the lowest count but has the biggest average of lifeExp at 80.7 while Afica Continent has the biggest count but the lowest average of lifeExp at 54.8
1. Oceania Continent has the biggest median of lifeExp at 52.9 while Afica Continent has the lowest at 52.9


###Measures of variability
##We wnat to know ‘How much is the data spread out from the middle?’
#Just looking at the data gives us a sense of this
#But we want break it down to one number so we can compare sample distributions

```{r}
# Look at the difference between each point and the mean
sum(x - mean(x))
```
##So we can square the differnce
#But this number will keep getting bigger as you add more observations
#We want something that is stable
```{r}
# Square each difference to get rid of negatives then sum
sum((x - mean(x))^2)
```
#Variance
```{r}
sum((x - mean(x))^2)/(length(x) - 1)
var(x)
```
#Standard Deviation
```{r}
sqrt(sum((x - mean(x))^2)/(length(x) - 1))
sd(x)
```
#Inter Quartile Range
```{r}
summary(x)
IQR(x)
```
#Range
```{r}
max(x)
min(x)
diff(range(x))
```
__EXPLANATION__:
1. The Variance of x is 1.36
2. The Standard Deviation is 1.17
3. The Inter Quartile Range is 0.5
4. The Maximum is 79
5. The Minimum is 75
6. The range is 4

##Calculate spread measures
```{r}
str(gap2007)
```
```{r}
# Compute groupwise measures of spread
gap2007 %>%
  group_by(continent) %>%
  summarize(sd(lifeExp),
            IQR(lifeExp),
            n())
```
```{r}
# Generate overlaid density plots
gap2007 %>%
  ggplot(aes(x = lifeExp, fill = continent)) +
  geom_density(alpha = 0.3)
```


#Choose measures for center and spread
```{r}
# Compute stats for lifeExp in Americas
head(gap2007)
```
```{r}
gap2007 %>%
  filter(continent == "Americas") %>%
  summarize(mean(lifeExp),
            sd(lifeExp))
```
__EXPLANATION__:
1. The average lifeExp from Americas Continent is 73.6 years

```{r}
# Compute stats for population
gap2007 %>%
  summarize(median(pop),
            IQR(pop))
```


##Shape and transformations
#4 chracteristics of a distribution that are of interest:
-center
 already covered
-spread or variablity
 already covered
-shape
 -modality: number of prominent humps (uni, bi, multi, or uniform - no humps)
 -skew (right, left, or symetric)
 -Can transform to fix skew
-outliers

#Transformations

```{r}
# Create density plot of old variable
gap2007 %>%
  ggplot(aes(x = pop)) +
  geom_density()
```

```{r}
# Transform the skewed pop variable
gap2007 <- gap2007 %>%
  mutate(log_pop = log(pop))

# Create density plot of new variable
gap2007 %>%
  ggplot(aes(x = log_pop)) +
  geom_density()
```
#Outliers
```{r}
# Filter for Asia, add column indicating outliers
str(gapminder)
```
```{r}
gap_asia <- gap2007 %>%
  filter(continent == "Asia") %>%
  mutate(is_outlier = lifeExp < 50)

# Remove outliers, create box plot of lifeExp
gap_asia %>%
  filter(!is_outlier) %>%
  ggplot(aes(x = 1, y = lifeExp)) +
  geom_boxplot()
```



###Case Study
-Spam and num_char

```{r}
email <- read.csv(file = 'email.csv')
```

```{r}
email %>%
  group_by(spam) %>%
  summarize( 
    median(num_char),
    IQR(num_char))
```
```{r}
str(email)
```

```{r}
table(email$spam)
```
```{r}
email <- email %>%
  mutate(spam = factor(ifelse(spam == 0, "not-spam", "spam")))

# Create plot
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = spam, y = log_num_char)) +
  geom_boxplot()
```
__EXPLANATION__:
1. There's is more not-spam than spam in email


```{r}
# Compute center and spread for exclaim_mess by spam
email %>%
  group_by(spam) %>%
  summarize(
    median(exclaim_mess),
    IQR(exclaim_mess))
```

```{r}
table(email$exclaim_mess)
```

```{r}
# Create plot for spam and exclaim_mess
email %>%
  mutate(log_exclaim_mess = log(exclaim_mess)) %>%
  ggplot(aes(x = log_exclaim_mess)) + 
  geom_histogram() + 
  facet_wrap(~ spam)
```
__EXPLANATION__:
1. The exclamation is more on the not-spam side rather than the spam side


```{r}
table(email$image)
```
```{r}
# Create plot of proportion of spam by image
email %>%
  mutate(has_image = image > 0) %>%
  ggplot(aes(x = has_image, fill = spam)) +
  geom_bar(position = "fill")
```
__EXPLANATION__:
1. The email that doesn't has image is more likely to get a spam


```{r}
# Test if images count as attachments
sum(email$image > email$attach)
```
##Answering questions with chains

```{r}
## Within non-spam emails, is the typical length of emails shorter for 
## those that were sent to multiple people?
email %>%
   filter(spam == "not-spam") %>%
   group_by(to_multiple) %>%
   summarize(median(num_char))
```
__EXPLANATION__:
1. Emails that sent to multiple people is usually shorter


```{r}
# Question 1
## For emails containing the word "dollar", does the typical spam email 
## contain a greater number of occurences of the word than the typical non-spam email?
table(email$dollar)
```

```{r}
email %>%
  filter(dollar > 0) %>%
  group_by(spam) %>%
  summarize(median(dollar))
```

__EXPLANATION__:
1. In the email that contain dollar, the not-spam is occur more often rather than the one who get spam

```{r}
# Question 2
## If you encounter an email with greater than 10 occurrences of the word "dollar", 
## is it more likely to be spam or not -spam?

email %>%
  filter(dollar > 10) %>%
  ggplot(aes(x = spam)) +
  geom_bar()
```
__EXPLANATION__:
1. The word dollar appear in a email that's not a spam


##What’s in a number?
```{r}
email$number <- as.factor(email$number)
levels(email$number)
```

```{r}
table(email$number)
```

```{r}
# Reorder levels
email$number <- factor(email$number, levels = c("none","small","big"))

# Construct plot of number
ggplot(email, aes(x = number)) +
  geom_bar() + 
  facet_wrap( ~ spam)
```

__EXPLANATION__:
1. In spam and not-spam, the number that appear the most is small number in both of them.
2. In a not-spam email, the least likely email to appear is an email that doesn't have number
3. In a spam email, the least likely email to appear is an email that has big number


